### 2023-09-06 proofs/main.tex
Add deterinicity lemmas, get rid of (??)

@@

Unification constraint entries are a special case of subtyping constraint
entries. They are merged using the same algorithm 
(\cref{definition:merge-matching-entries}).
Notice that the merge of two matching unification constraint entries
is a unification constraint entry.
\begin{lemma}[Merge of Matching Unification Constraint Entries is well-defined]
  Suppose that $[[Γ ⊢ ucE1]]$ and $[[Γ ⊢ ucE2]]$
  are unification constraint entries. 
  Then the merge of $[[ucE1]]$ and $[[ucE2]]$ 
  $[[Γ ⊢ ucE1 & ucE2 = ucE]]$
  according to \cref{definition:merge-matching-entries},
  is a unification constraint entry.
\end{lemma}
\begin{proof}
  Since $[[ucE1]]$ and $[[ucE2]]$ are matching unification constraint entries,
  they have the shape $([[pua :≈ iP1]], [[pua :≈ iP2]])$ or
  $([[nua :≈ iN1]], [[nua :≈ iN2]])$.
  Then the merge of $[[ucE1]]$ and $[[ucE2]]$ 
  can only be defined by \ruleref{\ottdruleSCMEPEqEqLabel} or
  \ruleref{\ottdruleSCMENEqEqLabel}.
  In both cases the result, if it exists, 
  is a unification constraint entry:
  in the first case, the result has shape $[[pua :≈ iP1]]$,
  in the second case, the result has shape $[[nua :≈ iN1]]$.
\end{proof}

@@

\begin{lemma}[Merge of Unification Constraints is well-defined]
  Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$
  are unification constraints. 
  Then the merge of $[[UC1]]$ and $[[UC2]]$ 
  $[[Θ ⊢ lift UC1 & lift UC2 = lift UC]]$
  according to \cref{definition:merge-subtyping-constraints},
  is a unification constraint.
\end{lemma}
\begin{proof}
  $[[UC]]$ consists of unmatched entries of $[[UC1]]$ and $[[UC2]]$,
  which are \emph{unification} constraint entries by assumption,
  and merge of matching entries, which also are  
  \emph{unification} constraint entries by \cref{lemma:merge-matching-entries-is-well-defined}.
\end{proof}

@@

\subsection{Anti-unification}
\input{../_build/_gen/tex/anti-unification-lemmas.tex}

>>>>

\subsection{Fresh Variable Selection}
\label{sec:fresh-selection}
Both the subtyping and the type inference algorithm
rely on the ability to select fresh, unused variables.
In the inference rules, it is indicated as 
$[[{varset} are fresh]]$. 
We assume that the selection subroutine always succeeds and is 
deterministic. In other words, whenever it is called in 
an algorithmic inference rule, it returns the same result, 
uniquely determined by the input of this rule.

@@

Unification constraint entries are a special case of subtyping constraint
entries. They are merged using the same algorithm 
(\cref{definition:merge-matching-entries}).
Notice that the merge of two matching unification constraint entries
is a unification constraint entry.
\begin{lemma}[Merge of Matching Unification Constraint Entries is well-defined]
  \label{lemma:merge-matching-entries-welldef}
  Suppose that $[[Γ ⊢ ucE1]]$ and $[[Γ ⊢ ucE2]]$
  are unification constraint entries. 
  Then the merge of $[[ucE1]]$ and $[[ucE2]]$ 
  $[[Γ ⊢ ucE1 & ucE2 = ucE]]$
  according to \cref{definition:merge-matching-entries},
  is a unification constraint entry.
\end{lemma}
\begin{proof}
  Since $[[ucE1]]$ and $[[ucE2]]$ are matching unification constraint entries,
  they have the shape $([[pua :≈ iP1]], [[pua :≈ iP2]])$ or
  $([[nua :≈ iN1]], [[nua :≈ iN2]])$.
  Then the merge of $[[ucE1]]$ and $[[ucE2]]$ 
  can only be defined by \ruleref{\ottdruleSCMEPEqEqLabel} or
  \ruleref{\ottdruleSCMENEqEqLabel}.
  In both cases the result, if it exists, 
  is a unification constraint entry:
  in the first case, the result has shape $[[pua :≈ iP1]]$,
  in the second case, the result has shape $[[nua :≈ iN1]]$.
\end{proof}

@@

\begin{lemma}[Merge of Unification Constraints is well-defined]
  Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$
  are unification constraints. 
  Then the merge of $[[UC1]]$ and $[[UC2]]$ 
  $[[Θ ⊢ lift UC1 & lift UC2 = lift UC]]$
  according to \cref{definition:merge-subtyping-constraints},
  is a unification constraint.
\end{lemma}
\begin{proof}
  $[[UC]]$ consists of unmatched entries of $[[UC1]]$ and $[[UC2]]$,
  which are \emph{unification} constraint entries by assumption,
  and merge of matching entries, which also are  
  \emph{unification} constraint entries by \cref{lemma:merge-matching-entries-welldef}.
\end{proof}

@@

\subsection{Anti-unification}
\input{../_build/_gen/tex/au-lemmas.tex}

### 2023-09-06 proofs/tex/alg-neg-subtyping-lemmas.tex
Add deterinicity lemmas, get rid of (??)

>>>>

\begin{observation}[Negative Algorithmic Subtyping is Deterministic]
    \label{obs:neg-subt-deterministic}
    For fixed $[[Γ]]$, $[[Θ]]$, $[[iM]]$, and $[[uN]]$, 
    if $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$ 
    and $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC']]$
    then $[[SC]]$ = $[[SC']]$.
\end{observation}
\begin{proof}
    First, notice that the shape of
    the input uniquely determines the rule applied to infer
    $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$,
    which is consequently, the same rule used to
    infer $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC']]$.

    Second, notice that for each of the inference rules, 
    the premises are deterministic on the input.
    Specifically, 
    \begin{itemize}
        \item \ruleref{\ottdruleAShiftULabel} relies on unification,
            which is deterministic by \cref{obs:unif-deterministic};
        \item \ruleref{\ottdruleAForallLabel} relies on
            the choice of fresh algorithmic variables,
            which is deterministic as discussed in \cref{sec:fresh-selection},
            and on the negative subtyping, which is deterministic by
            the induction hypothesis;
        \item \ruleref{\ottdruleAArrowLabel} uses 
            the negative subtyping 
            (deterministic by the induction hypothesis),
            the positive subtyping 
            (\cref{obs:pos-subt-deterministic}),
            and the merge of subtyping constraints
            (\cref{obs:subt-merge-deterministic});
    \end{itemize}
\end{proof}

### 2023-09-06 proofs/tex/alg-pos-subtyping-lemmas.tex
Add deterinicity lemmas, get rid of (??)

>>>>

\begin{observation}[Positive Subtyping is Deterministic]
    \label{obs:pos-subt-deterministic}
    For fixed $[[Γ]]$, $[[Θ]]$, $[[uP]]$, and $[[iQ]]$,
    if $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ and $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC']]$
    then $[[SC]] = [[SC']]$.
\end{observation}
\begin{proof}
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$.
    First, it is easy to see that the rule applied to infer 
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ uniquely depends on the input, 
    and those, it is the same rule that is inferring
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC']]$.
    Second, the premises of each rule are deterministic on the input: 
    unification is deterministic by \cref{obs:unif-deterministic},
    upgrade is deterministic by \cref{obs:upgrade-deterministic},
    the choice of the fresh algorithmic variables is deterministic by
    convention, as discussed in \cref{sec:fresh-selection},
    positive subtyping by the induction hypothesis.
\end{proof}
